### 2020-07-01 README.Rmd
update README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
figures_rel_path_from_root <- rprojroot::find_root_file(
  "man", "figures", "README", 
  criterion = rprojroot::has_file("DESCRIPTION"))

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = figures_rel_path_from_root,
  out.width = "100%"
)
```

# eurobarometer

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/antaldaniel/eurobarometer.svg?branch=master)](https://travis-ci.com/antaldaniel/eurobarometer)
[![Codecov test coverage](https://codecov.io/gh/antaldaniel/eurobarometer/branch/master/graph/badge.svg)](https://codecov.io/gh/antaldaniel/eurobarometer?branch=master)
[![Project Status: Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
"[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)"
"[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/eurobarometer)](https://cran.r-project.org/package=eurobarometer)"
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3901724.svg)](https://doi.org/10.5281/zenodo.3901724)
[![Follow author](https://img.shields.io/twitter/follow/antaldaniel.svg?style=social)](https://twitter.com/intent/follow?screen_name=antaldaniel)
[![Follow contributor](https://img.shields.io/twitter/follow/martakolcz.svg?style=social)](https://twitter.com/martakolcz?lang=en)
<!-- badges: end -->

The goal of `eurobarometer` is converting Eurobarometer microdata files, as stored by GESIS, into tidy R data frames and help common pre-processing problems.

Please report all issues to [github.com/antaldaniel/eurobarometer/issues](https://github.com/antaldaniel/eurobarometer/issues).

Pull requests are welcome on [github.com/antaldaniel/eurobarometer]((https://github.com/antaldaniel/eurobarometer/issues))  from all potential contributors who abide by  the terms of the [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).

If you use `eurobarometer` in your work, please [cite the package](https://doi.org/10.5281/zenodo.3825700).

## Installation

At this moment you cannot install yet the package from [CRAN](https://CRAN.R-project.org), but there is a development version on [GitHub](https://github.com/):

``` r
# install.packages("devtools")
devtools::install_github("antaldaniel/eurobarometer")
```

## Data Preparation For Programmatic Use

The SPSS files in GESIS are not ready for programmatic use in R or for joining into longitudinal panels.

1. We import the GESIS SPSS files with the help of the [haven](https://haven.tidyverse.org/) package. In order to preserve the metadata in the files (particularly variable- and value labels, and user-defined missing values) these are converted to labelled classes.

2. We analyse the files with `gesis_metadata_create()` to understand what are the possible problems in the SPSS file and how to resolve them. (See: [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html) or `vignette("metadata")`)

3. We suggest various [naming](http://eurobarometer.danielantal.eu/articles/variable_names.html) and [harmonization](http://eurobarometer.danielantal.eu/articles/workflow.html) changes in the data as a data processing step. These suggested changes can be modified by the user.

4. We convert the variables to a modified version of the `haven_labelled_spss` class that contains the harmonized values, harmonized labels, harmonized missing values codes of the variable with their history (original values, labels, and function applied to change them)  for reproducibility. The [eurobarometer_labelled](http://eurobarometer.danielantal.eu/articles/eurobarometer_class.html) class inherits most of the functionality of the packages [haven](https://haven.tidyverse.org/),  [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) and are expected to work well with [sjlabelled](https://strengejacke.github.io/sjlabelled/articles/labelleddata.html).

```{r, echo=FALSE, fig.align='center', out.width='90%', fig.cap="Workflow from 'Introduction to labelled by Joseph Larmaranged'"}
knitr::include_graphics(file.path(figures_rel_path_from_root, 
                                  'labelled_workflow.png'))
```

Our approach follows `Approach B` to allow the conversion of harmonized and joined files back into SPSS.

5. The [eurobarometer_labelled](http://eurobarometer.danielantal.eu/articles/eurobarometer_class.html) class comes with simple methods that can convert all variables to the base R types of numeric or factor, and they are ready to be used with any R statistical packages and tidyverse data processing tools. Alternatively, they can be exported back to [SPSS](https://haven.tidyverse.org/reference/read_spss.html) or [Stata](https://haven.tidyverse.org/reference/read_dta.html) files with the functions of the [haven](https://haven.tidyverse.org/) package.

## Joining With Eurostat & Other Data Tables

*  The regional boundaries are consistently used, coded and named - this will be harmonized with the package [regions](http://regions.danielantal.eu/), which helps resolving common issues, and to properly join various Eurobarometer files with Eurostat data tables and Google data tables.

* This will also allow you to create regional statistics from Eurobarometer microdata files.

## Code of Conduct

Please note that the `eurobarometer` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to its terms.