### 2020-07-03 data-raw/variable-keywords.Rmd
draft variable-keywords vignette

---
title: "Tagging variables with keywords"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vocabulary}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
options(width = 160)
```

```{r setup}
library(eurobarometer)
library(tidyverse)
library(tibble)
library(knitr)
library(kableExtra)
library(tidytext)
library(janitor)
```

An important aspect of survey data harmonization - in addition to standardizing response coding and value labels - is the identification of variables that measure the same thing, and standardizing the name of these variables. The process is faciliated by the fact that variables typically have labels that succintly describe the content of the variable, and sometimes also contain information about the coding of responses (e.g., length of scale), the unit (e.g., age in years or categories) and whether the variable contains "original" respondent responses or some transformation of responses (e.g. recoding or index).

Fortunately, in many multi-wave surveys, such as the Eurobarometer, there is a degree of consistency in how variables are labelled in different editions of the survey, and most survey project try to follow some labelling conventions.

Unfortunately, consistency is limited and conventions are not followed strictly, which makes standardizing labels a semi-manual task. Some of these issues stem from the limits in the number of characters of variable labels in common statistical software packages (e.g., SPSS), which leads to non-standard abbreviations.

This vignette outlines a procedure for arriving at standardized variable names via tokenization of normalized variable labels (`var_label_norm`), which are created by [`label_normalize`](http://eurobarometer.danielantal.eu/reference/label_normalize.html) as part of [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html). The logic is similar to text analysis, in that a document-term matrix is created with each variable label treated as a (very short) document, and terms treated as keywords. variables that contain the same sets of keywords can be - with some margin of error - assumed to correspond to the same or equivalent survey questions.

## Procedure

0. Read metadata table (created by [`gesis_metadata_create`](http://eurobarometer.danielantal.eu/reference/gesis_metadata_create.html)),

1. Standardize the spelling of certain often abbreviated words (e.g., 'gov', 'govnmt', and 'govmnt' for 'government'),

2. Tokenize,

3. Eliminate stop words (but be careful to not eliminate too many, e.g. 'right' or 'working),

4. Create a document-term matrix (actually, variable label - keyword matrix),

5. Sort the columns according to the declining frequency of the keywords,

6. Create standardized variable labels by concatenating keywords.

### 0. Metadata table

The table contains metadata from 97 Eurobarometer files (each file corresponding to one EB round).

```{r}
metadata_database <- readRDS( file.path('../data-raw', 'eb_metadata_database_20200628.rds'))
```

### 1. Standardize the spelling

This would perhaps look better as a translation table instead of regex.

Maybe the recoding of "leftover question numbers at the beginning of value labels" can become part of `label_normalize`?

```{r}

dat <- metadata_database %>%
  distinct(filename, qb, var_name_orig, var_name_orig, var_label_orig, var_label_norm) %>%

  # eliminate variables that look like region identifiers, weights, archive IDs, etc.
  filter(!grepl("recoded|(rec)|nuts|iso_3166|iso3166|region_ii", var_label_norm)) %>%
  filter(!grepl("^region_|sample_id|(sum)|(index)|filename", var_label_norm)) %>%
  filter(!grepl("icpsr|gesis|zacat|weight_", var_label_norm)) %>%

  # clean up variable labels to unify the spelling of certain common words and abbreviation
  mutate(var_id = paste0(filename, " ", var_name_orig),

         var_label_norm1 = gsub("^cap_", "common-agricultural-policy_", var_label_norm),
         var_label_norm1 = gsub("_cap_", "_common-agricultural-policy_", var_label_norm),

         # leftover question numbers at the beginning of value labels
         var_label_norm1 = gsub("^w[0-9]{1,2}", "", var_label_norm1),
         var_label_norm1 = gsub("^[0-9]{1-3}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]{2}[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z]{1,2}[0-9]{1,2}[a-z]_", "", var_label_norm1),
         var_label_norm1 = gsub("^[0-9]{1,2}_", "", var_label_norm1),
         var_label_norm1 = gsub("^[a-z][0-9]_[0-9]{1,2}_", "", var_label_norm1),

         # country names
         var_label_norm1 = gsub("united_kingdom", "united-kingdom", var_label_norm1),
         var_label_norm1 = gsub("czech_republic", "czech-republic", var_label_norm1),
         var_label_norm1 = gsub("great_britain", "great-britain", var_label_norm1),
         var_label_norm1 = gsub("united_germany", "united-germany", var_label_norm1),
         var_label_norm1 = gsub("united_nations", "united-nations", var_label_norm1),
         var_label_norm1 = gsub("ivory_coast", "ivory-coast", var_label_norm1),

         # commonly abbreviated words
         var_label_norm1 = gsub("particip_", "participation_", var_label_norm1),
         var_label_norm1 = gsub("environmt_|environm_", "environment_", var_label_norm1),
         var_label_norm1 = gsub("citizenshp_", "citizenship_", var_label_norm1),
         var_label_norm1 = gsub("_rght", "_right", var_label_norm1),
         var_label_norm1 = gsub("_fulfld", "_fullfilled", var_label_norm1),
         var_label_norm1 = gsub("_pessim", "_pessimistic", var_label_norm1),
         var_label_norm1 = gsub("_convcd", "_convinced", var_label_norm1),
         var_label_norm1 = gsub("not_invo", "not_involved", var_label_norm1),
         var_label_norm1 = gsub("intelligenc$", "intelligence", var_label_norm1),
         var_label_norm1 = gsub("gvrnm", "government", var_label_norm1),
         var_label_norm1 = gsub("ctry|cntry|\\[country\\]", "country", var_label_norm1),
         var_label_norm1 = gsub("_accom$|_accomod$", "_accomodation", var_label_norm1),
         var_label_norm1 = gsub("_activit", "_activity", var_label_norm1),
         var_label_norm1 = gsub("_cig|_cigs", "_cigarettes", var_label_norm1),
         var_label_norm1 = gsub("cildren", "children", var_label_norm1),
         var_label_norm1 = gsub("citizenshp", "citizenship", var_label_norm1),
         var_label_norm1 = gsub("collgs|collegues", "colleagues", var_label_norm1),
         var_label_norm1 = gsub("competitivn", "competition", var_label_norm1),
         var_label_norm1 = gsub("comptrs", "computers", var_label_norm1),
         var_label_norm1 = gsub("confidfence", "confidence", var_label_norm1),
         var_label_norm1 = gsub("conflicte$", "conflicts", var_label_norm1),
         var_label_norm1 = gsub("conn$|connec$", "connection", var_label_norm1),
         var_label_norm1 = gsub("contrls", "controls", var_label_norm1),
         var_label_norm1 = gsub("communit$", "community", var_label_norm1),
         var_label_norm1 = gsub("situatn_", "situation", var_label_norm1),
         var_label_norm1 = gsub("freq_", "frequency_", var_label_norm1),
         var_label_norm1 = gsub("hh", "household", var_label_norm1),
         var_label_norm1 = gsub("_pers_", "_personal_", var_label_norm1),
         var_label_norm1 = gsub("prov_", "provider_", var_label_norm1),
         var_label_norm1 = gsub("info_", "information_", var_label_norm1),
         var_label_norm1 = gsub("newspv|newspapers", "newspaper", var_label_norm1),
         var_label_norm1 = gsub("_newsp_|_newspap_", "_newspaper_", var_label_norm1),
         var_label_norm1 = gsub("_newsp$|_newspap$", "_newspaper_", var_label_norm1),
         var_label_norm1 = gsub("serv_", "services_", var_label_norm1),
         var_label_norm1 = gsub("financ_", "financial_", var_label_norm1),
         var_label_norm1 = gsub("_op_", "_operator_", var_label_norm1),

         # two- or three-word phrases to be kept as keywords
         var_label_norm1 = gsub("tv_channels", "tv-channels", var_label_norm1),
         var_label_norm1 = gsub("radio_stations", "radio-stations", var_label_norm1),
         var_label_norm1 = gsub("regional_local", "regional-local", var_label_norm1),
         var_label_norm1 = gsub("air_conditioning", "air-conditioning", var_label_norm1),
         var_label_norm1 = gsub("pol_party|political_party", "political-party", var_label_norm1),
         var_label_norm1 = gsub("electoral_particip_", "electoral-participation_", var_label_norm1),
         var_label_norm1 = gsub("electoral_particip$", "electoral-participation$", var_label_norm1),
         var_label_norm1 = gsub("trust_in_institutions", "trust", var_label_norm1),
         var_label_norm1 = gsub("justice_legal_system|justice_nat_legal_system", "justice-system", var_label_norm1),
         var_label_norm1 = gsub("non_govmnt_org|non_govnmt_org", "ngo", var_label_norm1),
         var_label_norm1 = gsub("polit_parties|political_parties", "political-parties", var_label_norm1),
         var_label_norm1 = gsub("reg_loc_public_authorities|reg_local_authorities|reg_loc_authorities|reg_local_public_authorities|rg_lc_authorities", "regional-local-authorities", var_label_norm1),
         var_label_norm1 = gsub("written_press", "press", var_label_norm1),
         var_label_norm1 = gsub("pers_influence", "personal-influence", var_label_norm1),
         var_label_norm1 = gsub("job_situation", "job-situation", var_label_norm1),
         var_label_norm1 = gsub("religious_inst$", "religious-institutions", var_label_norm1),
         var_label_norm1 = gsub("polit_matters|political_matters", "political-matters", var_label_norm1),
         var_label_norm1 = gsub("pol_discussion|polit_discussion|political_discussion", "political-discussion", var_label_norm1),
         var_label_norm1 = gsub("nat_resources", "natural-resources", var_label_norm1),
         var_label_norm1 = gsub("non_euro_zone", "non-euro-zone", var_label_norm1),
         var_label_norm1 = gsub("party_attachment", "party-attachment", var_label_norm1),
         var_label_norm1 = gsub("which_issue", "which-issue", var_label_norm1),
         var_label_norm1 = gsub("economic_situation|economic_sit", "economic-situation", var_label_norm1),
         var_label_norm1 = gsub("public_debt", "public-debt", var_label_norm1),
         var_label_norm1 = gsub("left_right", "left-right", var_label_norm1),
         var_label_norm1 = gsub("important_issues_cntry", "important-issues-cntry", var_label_norm1),
         var_label_norm1 = gsub("important_issues|import_issues", "important-issues", var_label_norm1),
         var_label_norm1 = gsub("important_values_eu", "important-values-eu", var_label_norm1),
         var_label_norm1 = gsub("important_values_pers", "important-values-personal", var_label_norm1),
         var_label_norm1 = gsub("important_life_domains", "important-life-domains", var_label_norm1),
         var_label_norm1 = gsub("quality_of_life|life_quality", "life-in-general", var_label_norm1),
         var_label_norm1 = gsub("household_composition", "household-composition", var_label_norm1),
         var_label_norm1 = gsub("size_of_community", "size-of-community", var_label_norm1),
         var_label_norm1 = gsub("head_of_hh", "head-of-hh", var_label_norm1),
         var_label_norm1 = gsub("last_job", "last-job", var_label_norm1),
         var_label_norm1 = gsub("level_in_society", "level-in-society", var_label_norm1),
         var_label_norm1 = gsub("self_placement", "self-placement", var_label_norm1),
         var_label_norm1 = gsub("social_class|soc_class", "social-class", var_label_norm1),
         var_label_norm1 = gsub("education_level", "education-level", var_label_norm1),
         var_label_norm1 = gsub("living_conditions", "living-conditions", var_label_norm1),
         var_label_norm1 = gsub("_nat_", "_national_", var_label_norm1),
         var_label_norm1 = gsub("^nat_", "national_", var_label_norm1),
         var_label_norm1 = gsub("_nat$", "_national", var_label_norm1),
         var_label_norm1 = gsub("europ_", "european_", var_label_norm1),
         var_label_norm1 = gsub("12_months_|12_mo_", "12-months_", var_label_norm1),
         var_label_norm1 = gsub("national_parliament", "national-parliament", var_label_norm1),
         var_label_norm1 = gsub("national_government", "national-government", var_label_norm1),
         var_label_norm1 = gsub("national_economy", "national-economy", var_label_norm1),
         var_label_norm1 = gsub("good_bad", "good-bad", var_label_norm1),
         var_label_norm1 = gsub("court_of_auditors", "court-of-auditors", var_label_norm1),
         var_label_norm1 = gsub("council_of_ministers", "council-of-ministers", var_label_norm1),

         # EU-related abbreviations and phrases
         var_label_norm1 = gsub("euro_12", "euro-12", var_label_norm1),
         var_label_norm1 = gsub("european_parliament|eu_parl", "european-parliament", var_label_norm1),
         var_label_norm1 = gsub("european_union", "european-union", var_label_norm1),
         var_label_norm1 = gsub("european_unification|europ_unification|europ_unif", "european-unification", var_label_norm1),
         var_label_norm1 = gsub("european_economy|europ_economy", "european-economy", var_label_norm1),
         var_label_norm1 = gsub("european_ombudsman", "european-ombudsman", var_label_norm1),
         var_label_norm1 = gsub("european_elections|ep_election", "european-elections", var_label_norm1),
         var_label_norm1 = gsub("european_elec|european_elect|europ_elec", "european-elections", var_label_norm1),
         var_label_norm1 = gsub("european_economy", "european-economy", var_label_norm1),
         var_label_norm1 = gsub("european_currency|europ_currency", "european-currency", var_label_norm1),
         var_label_norm1 = gsub("europ_court_of_auditors|eur_court_of_auditors", "eu-court-of-auditors", var_label_norm1),
         var_label_norm1 = gsub("europ_court_of_justice|european_court_of_justice", "eu-court-of-justice", var_label_norm1),
         var_label_norm1 = gsub("econ_and_social_committee|economic_and_soc_committee|econ_and_soc_committee", "econ-and-soc-committee", var_label_norm1),
         var_label_norm1 = gsub("europ_polit_matter|europ_political_matters", "european-political-matters", var_label_norm1),
         var_label_norm1 = gsub("euro_zone", "euro-zone", var_label_norm1),
         var_label_norm1 = gsub("eu_issues", "eu-issues", var_label_norm1),
         var_label_norm1 = gsub("eu_budget", "eu-budget", var_label_norm1),
         var_label_norm1 = gsub("eu_statements", "eu-statements", var_label_norm1),
         var_label_norm1 = gsub("eu_citizenship", "eu-citizenship", var_label_norm1),
         var_label_norm1 = gsub("eu_membership|eu_membersh", "eu-membership", var_label_norm1),
         var_label_norm1 = gsub("eu_image", "eu-image", var_label_norm1),
         var_label_norm1 = gsub("eu_level", "eu-level", var_label_norm1),
         var_label_norm1 = gsub("ecb|european_central_bank", "european-central-bank", var_label_norm1),
         var_label_norm1 = gsub("european_commission", "european-commission", var_label_norm1),
         var_label_norm1 = gsub("nms", "new-member-states", var_label_norm1),
         var_label_norm1 = gsub("council_of_the_eu", "council-of-the-eu", var_label_norm1),

         # replace underscore with space
         text = gsub("_", " ", var_label_norm1)) %>%
  ungroup() %>%
  # select the needed variables
  select(filename, qb, var_label_norm1, var_label_orig, var_id, text)
```

### 2 & 3. Tokenize & Eliminate stopwords

```{r}
# list of country codes to be converted to country names
eu_countries <- c("at", "be", "bg", "cy", "cz", "dk", "de", "ee", "es", "fr",
                  "fi", "gb", "gr", "hr", "hu", "ie", "it", "lt", "lv", "lu",
                  "mt", "nl", "pl", "pt", "ro", "se", "si", "sk", "tr")

# !!! the problem here is that dk sometimes means "don't know" and sometimes Demnark !!!

tidy <- dat %>%
  # tokenize
  unnest_tokens(word, text, token = "regex", pattern = " |  ") %>%
  # eliminate stop words, with some exceptions
  anti_join(stop_words %>% filter(!word %in% c("working", "right", "other"))) %>%
  # eliminate numbers
  filter(!grepl("\\b\\d+\\b", word)) %>%
  # convert country codes to country names
  mutate(country = ifelse(word %in% eu_countries, word, NA),
         country = countrycode::countrycode(toupper(country), "iso2c", "country.name"),
         country = tolower(country),
         word = ifelse(!is.na(country), country, word),
         word = ifelse(word == "united kingdom", "united-kingdom", word),
         word = ifelse(word == "tcc", "turkish-cypriot-community", word)) %>%
  select(-country)

head(tidy)
```

```{r}
tidy %>%
  count(word) %>%
  arrange(desc(n)) %>% 
  print(n = 50)
```

### 4. Create a document-term matrix

```{r}
# select words with > 5 occurrences
tidy_small <- tidy %>%
  group_by(word) %>%
  mutate(count_of_word = n()) %>%
  filter(count_of_word > 5)

dtm <- tidy_small %>%
  distinct(var_id, var_label_norm1, word) %>%
  mutate(n = word) %>%
  spread(word, n)

dtm[1100:1110, c(1,2,110:112)]
```

### 5. Sort columns according to the declining frequency of the keywords

```{r}
# get the order of variables according to declining proportion of missing value
# (declining popularity of keywords)
order <- dtm %>%
  summarise_each(funs(sum(is.na(.))/length(.))) %>%
  t() %>% as.vector()

names(order) <- names(dtm)
```

### 6. Create standardized variable labels by concatenating keywords

Standardized variable labels `var_label_std` are concatenated keywords, where the keywords occur in the order of their frequency in the entire corpus of variable labels. As a consequence, more common - likely broader - terms will come first, and less common - more specific - terms will come later.

```{r}
dtm2 <- dtm %>% select(names(sort(order))) %>%
  mutate_at(vars(3:ncol(dtm)), funs(ifelse(is.na(.), "", .))) %>%
  unite("var_label_std", 3:ncol(dtm)) %>%
  mutate(var_label_std = gsub("(\\_)\\1+", "\\1", var_label_std)) %>%
  mutate(var_label_std = gsub("^_|_$", "", var_label_std))

dtm2 %>% 
  print(n = 50)
```

Print the most common standardized variable labels.

```{r}
dtm2 %>%
  count(var_label_std) %>%
  # exclude nationality, country (nation) codes, size of community (country-specific), language of interview
  filter(!grepl("nationality|nation_|size-of-community|language_interview", var_label_std)) %>%
  filter(n > 5) %>%
  arrange(desc(n)) %>% 
  print(n = 50)
```

## Application: political attitudes (1)

Filter on substrings of standardized variable labels `var_label_std`, and count the repetitions across datasets.

```{r}
# filter on var_label_std
dtm2 %>%
  filter(grepl("trust|satisf|left-right|interest|class", var_label_std)) %>% 
  count(var_label_std) %>%
  arrange(desc(n)) %>%
  print(n = 50)
```

## Application: political attitudes (2)

Filter on keywords in the document-term matrix, e.g. `left-right` to see filenames and variables corresponding to variables measuring the respondent's left-right placement on the spectrum of political opinions. This is an easy variable, because all standardized variable labels look pretty much the same, only those with "harmonized" or the number of categories added would require a closer look - they are likely tranformed versions of original variables.

Another thing that is noticable is that variable names are standardized starting with [ZA5913: Eurobarometer 81.2 (March 2014)](https://dbk.gesis.org/dbksearch/sdesc2.asp?no=5913&db=e&doi=10.4232/1.12172), where all left-right scale variables have the name `d1`. Prior to that, each EB wave has its own variable name for the left-right scale variable.

Altogether there are 92 such variables, and eliminating the transformed ones (harmonized or recoded), we are left with 85 variables in 85 different EB waves.

```{r}
# search document-term matrix by keyword
dtm %>%
  filter(!is.na(`left-right`)) %>%
  janitor::remove_empty(c("rows", "cols")) %>% 
  print(n = 100)
```