### 2020-07-20 README.Rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
figures_rel_path_from_root <- rprojroot::find_root_file(
  "man", "figures", "README", 
  criterion = rprojroot::has_file("DESCRIPTION"))

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = figures_rel_path_from_root,
  out.width = "100%"
)
```

# eurobarometer

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/antaldaniel/eurobarometer.svg?branch=master)](https://travis-ci.com/antaldaniel/eurobarometer)
[![Codecov test coverage](https://codecov.io/gh/antaldaniel/eurobarometer/branch/master/graph/badge.svg)](https://codecov.io/gh/antaldaniel/eurobarometer?branch=master)
[![Project Status: Minimal or no implementation has been done yet, or the repository is only intended to be a limited example, demo, or proof-of-concept.](https://www.repostatus.org/badges/latest/concept.svg)](https://www.repostatus.org/#concept)
[![license](https://img.shields.io/badge/license-GPL--3-blue.svg)](https://www.gnu.org/licenses/gpl-3.0.en.html)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/eurobarometer)](https://cran.r-project.org/package=eurobarometer)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3901724.svg)](https://doi.org/10.5281/zenodo.3901724)
[![Follow author](https://img.shields.io/twitter/follow/antaldaniel.svg?style=social)](https://twitter.com/intent/follow?screen_name=antaldaniel)
[![Follow contributor](https://img.shields.io/twitter/follow/martakolcz.svg?style=social)](https://twitter.com/martakolcz?lang=en)
<!-- badges: end -->

The goal of `eurobarometer` is converting Eurobarometer microdata files, as stored by GESIS, into tidy R data frames and help common pre-processing problems. The more generic function for retrospecitve (ex post) survey harmonization can be found in the package [retroharmonize](http://retroharmonize.satellitereport.com/).

Please report all issues to [github.com/antaldaniel/eurobarometer/issues](https://github.com/antaldaniel/eurobarometer/issues).

Pull requests are welcome on [github.com/antaldaniel/eurobarometer]((https://github.com/antaldaniel/eurobarometer/issues))  from all potential contributors who abide by  the terms of the [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html).

If you use `eurobarometer` in your work, please [cite the package](https://doi.org/10.5281/zenodo.3825700).

## Installation

At this moment you cannot install yet the package from [CRAN](https://CRAN.R-project.org), but there is a development version on [GitHub](https://github.com/):

``` r
# install.packages("devtools")
devtools::install_github("antaldaniel/retroharmonize")
devtools::install_github("antaldaniel/eurobarometer")
```

You do not need to load the [retroharmonize](http://retroharmonize.satellitereport.com/) package with `library(retroharmonize)` but eurobarometer relies on some if its classes, methods and functions.

## Data Preparation For Programmatic Use

The microdata of the European Commissions Eurobarometer surveys are stored and accessible in [GESIS](https://www.gesis.org/home). They are not ready for programmatic use in R or for joining into longitudinal panels. We created this package to help this procedure following the principles of reproducible research.

1. We import the GESIS SPSS files with the help of the adopted version of `read_spss()` from the [haven](https://haven.tidyverse.org/) package. The `haven::write_sav()` function does not work perfectly, it loses some metadata, so the best is to save you data as an `.rds` file. You can read surveys then with `read_rds()`. Both importing function will try to add important metadata to the `retroharmonize::survey()` tibble/data.frame class, including the original file name, the survey's ID and DOI.

2. We analyse the files with `gesis_metadata_create()` to understand what are the possible problems in the SPSS file and how to resolve them. (See: [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html) or `vignette("metadata")`) This is a modified version of the more generic [metadata_create](http://retroharmonize.satellitereport.com/reference/metadata_create.html) function in `retroharmonize`. See the [Working With Metadata](http://eurobarometer.danielantal.eu/articles/metadata.html) vignette for some more details.

3. We suggest various [naming](http://eurobarometer.danielantal.eu/articles/variable_names.html) and [harmonization](http://eurobarometer.danielantal.eu/articles/workflow.html) changes in the data as a data processing step. These suggested changes can be modified by the user.

4. We convert the variables to a modified version of the `retroharmonize_labelled_spss_survey` class that contains the harmonized values, harmonized labels, harmonized missing values codes of the variable with their history (original values, labels, and function applied to change them)  for reproducibility. The [labelled_spss_survey](http://retroharmonize.satellitereport.com/articles/labelled_spss_survey.html) class inherits most of the functionality of the packages [haven](https://haven.tidyverse.org/),  [labelled](https://cran.r-project.org/web/packages/labelled/vignettes/intro_labelled.html) and are expected to work well with [sjlabelled](https://strengejacke.github.io/sjlabelled/articles/labelleddata.html).

5. The harmonization of the variable names is currently possible with retroharmonize only (See: [Harmonize Value Labels](http://retroharmonize.satellitereport.com/articles/harmonize_labels.html)). Further, Eurobarometer specific helpers will be added soon. See the article [Survey Data Harmonization](http://eurobarometer.danielantal.eu/articles/harmonization.html) as a case study.

```{r, echo=FALSE, fig.align='center', out.width='90%', fig.cap="Workflow from 'Introduction to labelled by Joseph Larmaranged'"}
knitr::include_graphics(file.path(figures_rel_path_from_root, 
                                  'labelled_workflow.png'))
```

Our approach follows `Approach B` to allow the conversion of harmonized and joined files back into SPSS.

## Joining With Eurostat & Other Data Tables

*  The regional boundaries are consistently used, coded and named - this will be harmonized with the package [regions](http://regions.danielantal.eu/), which helps resolving common issues, and to properly join various Eurobarometer files with Eurostat data tables and Google data tables.

* This will also allow you to create regional statistics from Eurobarometer microdata files.

## Code of Conduct

Please note that the `eurobarometer` project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to its terms.

## Disclaimer

The authors of this package are not affiliated with [GESIS](https://www.gesis.org/home), or the producer of the Eurobarometer surveys, or the European Commission. For more information about the European Commission's Eurobarometer history, visit the European Commission's [Public Opinion](https://ec.europa.eu/commfrontoffice/publicopinion/index.cfm) website.